---
title: 'Investigation #2: A Real (Computing) Cluster'
author: "Keerthana Bellary, Cole Rowland"
format: pdf
echo: true
---

```{r}
#| include: false
options(scipen=999)  # turns off scientific notation
set.seed(123456)  # set your own seed!
```

A computing cluster is shared by 250 independent subscribers. Each day, the probability any particular subscriber uses the cluster is 0.3. The number of tasks submitted by a user to the cluster can be modeled by a geometric distribution with a mean of 6.5. The time that it takes the cluster to complete a task can be modeled by an exponential distribution with mean 2.8 minutes. Completion times are also independent.

------------------------------------------------------------------------

Before you write any code, let's reason out some values that you can use later to validate your simulation. These are arithmetic questions; no R code is required.

### 1. What is the expected number of tasks submitted in a day?

<!-- Write your arithmetic here. -->

250*0.3*6.5 = 487.5

### 2. What is the expected amount of time required to complete all the day's tasks?

<!-- Write your arithmetic here. -->

487.5\*2.8 = 1365

------------------------------------------------------------------------

Now you will write code to simulate one day’s worth of tasks, in order to answer the questions below. Notice that the model has three random variables, and they're *nested*: the number of tasks depends on the number of users that day, and the time to complete all tasks depends on the total number of tasks submitted.

Your main function, `cluster()`, will have no inputs; you may hard-code the parameters 250, 0.3, 6.5, and 2.8. Your function must return the values of **two** variables: *X = the total number of tasks submitted*, and *Y = the total amount of time required to complete all tasks*.

After that, the output of your Repeat step should consist of two sets of numbers: the simulated values for *X* and the simulated values for *Y*.

Don't worry about unfinished tasks “bleeding over” into the next day; you are simulating independent days. Equivalently, assume each day starts fresh, with zero tasks queued up.

Two very important notes:

- Everything in `cluster()` must be **vectorized**. There should be no *for* loops in your code.

- R's parameterization of the geometric distribution is different. You will need to adjust for it accordingly. You have been warned.

------------------------------------------------------------------------

<!-- Place your Build function and Repeat code in the code chunk below. -->

```{r Build and Repeat Steps}
# Build
cluster <- function(){
users <- rbinom(1, size = 250, prob = 0.3)
p <- 1/(6.5+1) 
tasks_per_user <- rgeom(users, prob = p) + 1
X <- sum(tasks_per_user)
times <- rexp(X, rate = 1/2.8) 
Y <- sum(times)
return(c(X,Y))
}

# Repeat
n <- 100000
sims <- replicate(n, cluster())
X <- sims[1, ]
Y <- sims[2, ]
```

------------------------------------------------------------------------

Use the output of the previous code chunk to answer the following questions. You should only run that code chunk once.

### 3. Create a well-labeled histogram of 100,000 values of the variable *X = number of tasks submitted in a day*. Summarize the distribution (shape, center, spread) in complete sentences.

```{r Histogram of X}
hist(X, 
xlab = "Number of Tasks", 
ylab = "Frequency", 
main = "Tasks Submitted in One Day")
```

<!-- Your written summary goes below this line. -->

This histogram has a normal distribution which can be seen by its singular peak in the center and approximately symmetrical spread, forming a bell shape curve. Of the 100,000 values of the variable X = number of tasks submitted in a day, the most common outcome (the center) is the range of 550-600 tasks being submitted in one day, with a frequency of approximately 23,000 values falling into that range. The number of tasks completed in a day falls almost entirely between 300 and 900 for each generated value.

### 4. Calculate and interpret a 95% confidence interval for the probability that more than 500 tasks are submitted in a day.

```{r CI Prob 1}
over_500 <- as.numeric(X > 500) 
t.test(over_500, conf.level = 0.95)
```

The 95% confidence interval for there being more than 500 tasks submitted in a day is \[0.7699017, 0.7750983\]. We are 95% confident that on any given day, the probability of the number of tasks submitted being over 500 is between 76.99% and 77.51%.

### 5. Calculate and interpret a 95% confidence interval for the mean number of tasks submitted in a day. Also comment on whether this interval is consistent with the common-sense answer from earlier.

```{r CI mu_X}
t.test(X, conf.level = 0.95) 
```

<!-- Your written summary goes below this line. -->

The 95% confidence interval for the mean number of tasks submitted in a day is \[561.8565, 562.8640\]. We are 95% confident that on any given day, the mean number of tasks submitted will be between 561.8565 and 562.8640. We calculated the expected value of tasks submitted in a day to be 487.5, which is inconsistent with this confidence interval based on the simulation.

### 6. Calculate a 95% confidence interval for the standard deviation of the number of tasks submitted in a day.

When $n$ is very large and the distribution of $X$ is approximately normal, a 95% CI for the variance $\sigma^2_X$ has endpoints

$$\frac{s^2}{1 \pm 1.96 \sqrt{2/(n-1)}}$$ where $s$ is the sample standard deviation. A 95% CI for $\sigma_X$ results from taking the square root of the two endpoints.

You must **write a function** that takes in a numerical vector and outputs the two endpoints for the 95% CI for $\sigma_X$. Execute your function to print out the CI (inside the code chunk is fine). No interpretation required.

```{r CI sigma_X}
sigma_ci <- function(x){
  n <- length(x)
  s <- sd(x)
  
  lower <- (s^2)/(1+1.96*sqrt(2/(n-1)))
  upper <- (s^2)/(1-1.96*sqrt(2/(n-1)))
  
  return(sqrt(c(lower, upper)))
}

sigma_ci(X)
```

<!-- Your written summary goes below this line. -->

The 95% confidence interval for the standard deviation of the number of tasks submitted in a day is \[80.92526, 81.63774\].

### 7. Calculate and interpret a 95% prediction interval for a single future value of the number of tasks submitted in a day.

When the distribution of $X$ is approximately normal, a 95% PI is given by $$\bar{x} \pm 1.96 \cdot s \cdot \sqrt{1 + \frac{1}{n}}$$ This formula is superior to (more reliable than) the quantile method used in the previous investigation, but it's only valid for normal distributions.

You must **write a function** that takes in a numerical vector and outputs the two endpoints for this 95% PI. Make sure you include an interpretation!

```{r PI X}
pred_int <- function(x){
  n <- length(x)
  xbar <- mean(x)
  s <- sd(x)
  lower <- xbar-1.96*s*sqrt(1+1/n)
  upper <- xbar+1.96*s*sqrt(1+1/n)
  return(c(lower, upper))
}

pred_int(X)
```

<!-- Your written summary goes below this line. -->

The 95% prediction interval for a single future value of the number of tasks submitted in a day is \[403.0523, 721.6682\]. We are 95% confident that a single future value of the number of tasks submitted in a day will be between 403.0523 and 721.6682.

------------------------------------------------------------------------

### 8. Create a well-labeled histogram of 100,000 values of the variable *Y = time to complete a day’s worth of tasks*. Summarize the distribution (shape, center, spread) in complete sentences.

```{r Histogram of Y}
hist(Y, 
xlab = "Total Time (minutes)",
ylab = "Frequency", 
main = "Time to Complete Tasks in One Day") 
```

<!-- Your written summary goes below this line. -->

This histogram has a normal distribution which can be seen by its singular peak in the center and approximately symmetrical spread, forming a bell shape curve. Of the 100,000 values of the variable Y = time to complete a day's worth of tasks, the most common outcome (the center) is the range of 1500 to 1600 minutes, with a frequency of approximately 17,000 values falling into that range. The total time required to complete tasks in a day falls almost entirely between 800 and 2500 for each generated value.

### 9. Calculate and interpret a 95% confidence interval for the probability that a day’s worth of tasks can be completed within 24 hours.

```{r CI Prob 2}
within_24 <- as.numeric(Y < 1440) 
t.test(within_24, conf.level = 0.95) 
```

<!-- Your written summary goes below this line. -->

The 95% confidence interval for the probability that a day's worth of tasks can be completed in 24 hours is \[0.2901691, 0.2958109\]. We are 95% confident that on any given day, the probability that a day's worth of tasks can be completed in 24 hours is between 29.02% and 29.58%.

### 10. Calculate and interpret a 95% confidence interval for the mean completion time for a day’s worth of tasks. Also comment on whether this interval is consistent with the common-sense answer from earlier.

```{r CI mu_Y}
t.test(Y, conf.level = 0.95) 
```

<!-- Your written summary goes below this line. -->

The 95% confidence interval for the mean completion time for a day’s worth of tasks is \[1572.775, 1575.713\]. We are 95% confident that on any given day, the mean completion time for a day’s worth of tasks will be between 1572.775 and 1575.713 minutes. We calculated the expected amount of time required to complete all the day's tasks to be 1365 minutes, which is inconsistent with this confidence interval based on the simulation.

### 11. Calculate a 95% confidence interval for the standard deviation of the time to complete a day’s worth of tasks.

Use your earlier function to perform the calculation. No interpretation required.

```{r}
sigma_ci(Y)
```

The 95% confidence interval for the standard deviation of the completion time for a day’s worth of tasks is \[235.9722, 238.0497\].

### 12. Calculate and interpret a 95% prediction interval for a single future value of the time to complete a day’s worth of tasks.

Use your earlier function to perform the calculation. Don't forget the interpretation!

```{r}
pred_int(Y) 
```

<!-- Your written summary goes below this line. -->

The 95% prediction interval for a single future value of the completion time for a day’s worth of tasks is \[1109.713, 2038.774\]. We are 95% confident that a single future value of the completion time for a day’s worth of tasks will be between 1109.713 and 2038.774 minutes.

------------------------------------------------------------------------

Finally, since you've simulated *two* variables, it makes sense to explore the *relationship* between them.

### 13. Create a well-labeled *scatterplot* of the two variables. Describe what you see (direction, form, strength) in complete sentences, and explain why this makes sense.

```{r Scatterplot}
plot(X, Y,
     xlab = "Number of Tasks",
     ylab = "Total Time (minutes)",
     main = "Tasks vs Completion Time")
```

<!-- Your written summary goes below this line. -->

We see a strong, positive, linear relationship between number of tasks submitted in a day and the completion time for a day’s worth of tasks. This makes sense because these variables are logically super closely linked together. It is not hard to understand that, in general, submitting more tasks will take more time to complete, and submitting less tasks will take less time to complete.

### 14. Use your simulation to calculate a 95% confidence interval for the probability that more than 500 tasks are submitted in a day *and* that those tasks can be completed within 24 hours.

Note: that's a single event, not two separate questions. No interpretation required.

```{r CI Prob 3}
event <- as.numeric(X > 500 & Y < 1440) 
t.test(event, conf.level = 0.95) 
```

The 95% confidence interval for the probability that more than 500 tasks are submitted in a day and that they can be completed in 24 hours is \[0.07734795, 0.08069205\].
